\textbf{SPECFEM3D Cartesian}\\ \textbf{User Manual}

### 2015-06-22 doc/USER_MANUAL/03_mesh_generation.tex
Buildbot ran the code beautifier.
renamed check_mesh_quality_CUBIT_Abaqus.f90 to check_mesh_quality.f90

For an elastic material:
\begin{description}
\item [{material\_id}] An integer value which is unique for this material.
\item [{Vp}] P-wave speed of the material (given in m/s).
\item [{Vs}] S-wave speed of the material (given in m/s).
\item [{rho}] density of the material (given in kg/m$^{3}$).
\item [{Q}] quality factor to use in case of a simulation with attenuation
turned on. It should be between 1 and 9000. In case no attenuation
information is available, it can be set to zero. Please note that
your Vp- and Vs-speeds are given for a reference frequency. To change
this reference frequency, you change the value of \texttt{ATTENUATION\_f0\_REFERENCE}
in the main constants file \texttt{constants.h} found in subdirectory
\texttt{src/shared/}.
\item [{anisotropic\_flag}] Flag describing the anisotropic model to use
in case an anisotropic simulation should be conducted. See the file
\texttt{model\_aniso.f90} in subdirectory \texttt{src/generate\_databases/}
for an implementation of the anisotropic models. In case no anisotropy
is available, it can be set to zero.
\end{description}

@@

In directory "CUBIT\_GEOCUBIT" we provide a script that can help doing the above tasks of exporting a CUBIT mesh
to SPECFEM3D format automatically for you: "run\_cubit2specfem3d.py".
Just edit them to indicate the path to your local installation of CUBIT and also the name of the *.cub existing CUBIT mesh file
that you want to export to SPECFEM3D format. These scripts will do the conversion for you automatically except assigning material properties
to the different mesh layers. To do so, you will then need to edit the file
called "nummaterial\_velocity\_file" that will have just been created and change it from the prototype created:

{\footnotesize
\begin{verbatim}
0 1 vol1 --> syntax: #material_domain_id #material_id #rho #vp #vs #Q_kappa #Q_mu #anisotropy
0 2 vol2 --> syntax: #material_domain_id #material_id #rho #vp #vs #Q_kappa #Q_mu #anisotropy
\end{verbatim}
}

(where "vol1" and "vol2" here represent the volume labels that you have set while creating the mesh in CUBIT)
to for instance

@@

\noindent
Then, in the main directory, type
{\small
\begin{verbatim}
make xcheck_mesh_quality_CUBIT_Abaqus
\end{verbatim}
}
and use
{\small
\begin{verbatim}
./bin/xcheck_mesh_quality_CUBIT_Abaqus
\end{verbatim}
}
to generate an %not supported: AVS output file (\texttt{\small AVS\_meshquality.inp} in AVS UCD format) or
OpenDX output file (\texttt{\small DX\_mesh\_quality.dx}{\small )
that can be used to investigate mesh quality, e.g. skewness of elements,
and a Gnuplot histogram (}\texttt{\small mesh\_quality\_histogram.txt}{\small )
that can be plotted with gnuplot (type `}\texttt{\small gnuplot plot\_mesh\_quality\_histogram.gnu}{\small ').
The histogram is also printed to the screen. Analyze that skewness
histogram of mesh elements to make sure no element has a skewness
above approximately 0.75, otherwise the mesh is of poor quality (and
if even a single element has a skewness value above 0.80, then you
must definitely improve the mesh). If you want to start designing
your own meshes, this tool is useful for viewing your creations. You
are striving for meshes with elements with `cube-like' dimensions,
e.g., the mesh should contain no very elongated or skewed elements.}{\small \par}

@@

An efficient partitioning is important because it leverages the overall
running time of the application. It amounts to balance the number
of elements in each slice while minimizing the communication costs
resulting from the placement of adjacent elements on different processors.
\texttt{decompose\_mesh} depends on the SCOTCH library \citep{PeRo96},
which provides efficent static mapping, graph and mesh partitioning
routines. SCOTCH is a free software package developed by Fran\c{c}ois
Pellegrini et al. from LaBRI and INRIA in Bordeaux, France, downloadable
from the web page \url{https://gforge.inria.fr/projects/scotch/}.\\

@@

\noindent
where
\begin{itemize}
\item \texttt{Q} : quality factor (0=no attenuation)
\item \texttt{anisotropy\_flag} : 0=no anisotropy / 1,2,.. check with implementation
in \texttt{aniso\_model.f90}
\item \texttt{domain\_id} : 1=acoustic / 2=elastic
\end{itemize}
\item [{\texttt{NREGIONS}}] The number of regions in the mesh. In the following
lines, because the mesh is regular or 'almost regular', each region
is defined as:
\begin{verbatim}
NEX_XI_BEGIN NEX_XI_END NEX_ETA_BEGIN NEX_ETA_END NZ_BEGIN NZ_END material_ID
\end{verbatim}
\end{description}
%
The \texttt{INTERFACES\_FILE} parameter of \texttt{Mesh\_Par\_File}
defines the file which contains the settings of the topography grid
and of the interfaces grids. Topography is defined as a set of elevation
values on a regular 2D grid. It is also possible to define interfaces
between the layers of the model in the same way. The file needs to
define several parameters:
\begin{itemize}
\item The number of interfaces, including the topography. This needs to
be set at the first line. Then, from the bottom to the top of the
model, you need to define the grids with:
\item \texttt{SUPPRESS\_UTM\_PROJECTION} flag as described previously,
\item number of points along $x$ and $y$ direction (NXI and NETA),
\item minimal $x$ and $y$ coordinates (LONG\_MIN and LAT\_MIN),
\item spacing between points along $x$ and $y$ (SPACING\_XI and SPACING\_ETA)
and
\item the name of the file which contains the elevation values (in $y$.$x$
increasing order).
\end{itemize}
%
At the end of this file, you simply need to set the number of spectral
elements in the vertical direction for each layer. We provide a few
models in the {\texttt{EXAMPLES/}} directory.\\

>>>>

For an elastic material:
\begin{description}
\item [{material\_id}] An integer value which is unique for this material.
\item [{Vp}] P-wave speed of the material (given in m/s).
\item [{Vs}] S-wave speed of the material (given in m/s).
\item [{rho}] density of the material (given in kg/m$^{3}$).
\item [{Q}] quality factor to use in case of a simulation with attenuation
turned on. It should be between 1 and 9000. In case no attenuation
information is available, it can be set to zero. You can either specify a single Q value,
in which case it will be assumed to be pure shear attenuation $Q_{\mu}$, or two
separate values for bulk and shear attenuation, $Q_{\kappa}$ and $Q_{\mu}$ respectively.

Please note that your Vp- and Vs-speeds are given for a reference frequency. To change
this reference frequency, you change the value of \texttt{ATTENUATION\_f0\_REFERENCE}
in the main constants file \texttt{constants.h} found in subdirectory
\texttt{src/shared/}.
\item [{anisotropic\_flag}] Flag describing the anisotropic model to use
in case an anisotropic simulation should be conducted. See the file
\texttt{model\_aniso.f90} in subdirectory \texttt{src/generate\_databases/}
for an implementation of the anisotropic models. In case no anisotropy
is available, it can be set to zero.
\end{description}

@@

In directory "CUBIT\_GEOCUBIT/" we provide a script that can help doing the above tasks of exporting a CUBIT mesh
to SPECFEM3D format automatically for you: "run\_cubit2specfem3d.py".
Just edit them to indicate the path to your local installation of CUBIT and also the name of the *.cub existing CUBIT mesh file
that you want to export to SPECFEM3D format. These scripts will do the conversion for you automatically except assigning material properties
to the different mesh layers. To do so, you will then need to edit the file
called "nummaterial\_velocity\_file" that will have just been created and change it from the prototype created:

{\footnotesize
\begin{verbatim}
0 1 vol1 --> syntax: #material_domain_id #material_id #rho #vp #vs #Q_kappa #Q_mu #anisotropy
0 2 vol2 --> syntax: #material_domain_id #material_id #rho #vp #vs #Q_kappa #Q_mu #anisotropy
\end{verbatim}
}
\noindent
(where "vol1" and "vol2" here represent the volume labels that you have set while creating the mesh in CUBIT)
to for instance

@@

\noindent
Then, in the main directory, type
{\small
\begin{verbatim}
make xcheck_mesh_quality
\end{verbatim}
}
and use
{\small
\begin{verbatim}
./bin/xcheck_mesh_quality
\end{verbatim}
}
to generate an %not supported: AVS output file (\texttt{\small AVS\_meshquality.inp} in AVS UCD format) or
OpenDX output file (\texttt{\small DX\_mesh\_quality.dx}{\small )
that can be used to investigate mesh quality, e.g. skewness of elements,
and a Gnuplot histogram (}\texttt{\small mesh\_quality\_histogram.txt}{\small )
that can be plotted with gnuplot (type `}\texttt{\small gnuplot plot\_mesh\_quality\_histogram.gnu}{\small ').
The histogram is also printed to the screen. Analyze that skewness
histogram of mesh elements to make sure no element has a skewness
above approximately 0.75, otherwise the mesh is of poor quality (and
if even a single element has a skewness value above 0.80, then you
must definitely improve the mesh). If you want to start designing
your own meshes, this tool is useful for viewing your creations. You
are striving for meshes with elements with `cube-like' dimensions,
e.g., the mesh should contain no very elongated or skewed elements.}{\small \par}

@@

An efficient partitioning is important because it leverages the overall
running time of the application. It amounts to balance the number
of elements in each slice while minimizing the communication costs
resulting from the placement of adjacent elements on different processors.
\texttt{decompose\_mesh} depends on the SCOTCH library \citep{PeRo96},
which provides efficient static mapping, graph and mesh partitioning
routines. SCOTCH is a free software package developed by Fran\c{c}ois
Pellegrini et al. from LaBRI and INRIA in Bordeaux, France, downloadable
from the web page \url{https://gforge.inria.fr/projects/scotch/}.\\

@@

\noindent
where
\begin{itemize}
\item \texttt{Q} : quality factor (0=no attenuation) for shear attenuation $Q_{\mu}$
\item \texttt{anisotropy\_flag} : 0=no anisotropy / 1,2,.. check with implementation
in \texttt{aniso\_model.f90}
\item \texttt{domain\_id} : 1=acoustic / 2=elastic
\end{itemize}
\item [{\texttt{NREGIONS}}] The number of regions in the mesh. In the following
lines, because the mesh is regular or 'almost regular', each region
is defined as:
\begin{verbatim}
NEX_XI_BEGIN NEX_XI_END NEX_ETA_BEGIN NEX_ETA_END NZ_BEGIN NZ_END material_ID
\end{verbatim}
\end{description}
%
The \texttt{INTERFACES\_FILE} parameter of \texttt{Mesh\_Par\_File}
defines the file which contains the settings of the topography grid
and of the interfaces grids. Topography is defined as a set of elevation
values on a regular 2D grid. It is also possible to define interfaces
between the layers of the model in the same way. The file needs to
define several parameters:
\begin{itemize}
\item The number of interfaces, including the topography. This needs to
be set at the first line. Then, from the bottom to the top of the
model, you need to define the grids with:
\item \texttt{SUPPRESS\_UTM\_PROJECTION} flag as described previously,
\item number of points along $x$ and $y$ direction (NXI and NETA),
\item minimal $x$ and $y$ coordinates (LONG\_MIN and LAT\_MIN),
\item spacing between points along $x$ and $y$ (SPACING\_XI and SPACING\_ETA)
and
\item the name of the file which contains the elevation values (in $y$.$x$
increasing order).
\end{itemize}
%
At the end of this file, you simply need to set the number of spectral
elements in the vertical direction for each layer. We provide a few
models in the {\texttt{EXAMPLES/}} directory.\\

### 2015-06-22 doc/USER_MANUAL/07_adjoint_simulations.tex

\begin{enumerate}
\item First, run a regular forward simlation (\texttt{SIMULATION\_TYPE =
1} and \texttt{SAVE\_FORWARD = .false.}). You can automatically set
these two variables using the \texttt{\small utils/change\_simulation\_type.pl}{\small{}
script:}{\small \par}
\begin{verbatim}
utils/change_simulation_type.pl -f
\end{verbatim}

@@

If \texttt{ANISOTROPIC\_KL} \texttt{=} \texttt{.true.} by additionally
setting \texttt{ANISOTROPIC\_KL} \texttt{=} \texttt{.true.} in \texttt{constants.h}
the package will save anisotropic kernels paramterized as velocities
related to transverse isotropy based on the the Chen and Tromp parameters
\citet{ChTr07}. The kernels are saved as relative perturbations for
horizontal and vertical P and S velocities, $\alpha_{v},\alpha_{h},\beta_{v},\beta_{h}$.
Explicit relations can be found in appendix B. of \citet{SiLiTrTr07b}

>>>>

\begin{enumerate}
\item First, run a regular forward simulation (\texttt{SIMULATION\_TYPE =
1} and \texttt{SAVE\_FORWARD = .false.}). You can automatically set
these two variables using the \texttt{\small utils/change\_simulation\_type.pl}{\small{}
script:}{\small \par}
\begin{verbatim}
utils/change_simulation_type.pl -f
\end{verbatim}

@@

If \texttt{ANISOTROPIC\_KL} \texttt{=} \texttt{.true.} by additionally
setting \texttt{ANISOTROPIC\_KL} \texttt{=} \texttt{.true.} in \texttt{constants.h}
the package will save anisotropic kernels parameterized as velocities
related to transverse isotropy based on the the Chen and Tromp parameters
\citet{ChTr07}. The kernels are saved as relative perturbations for
horizontal and vertical P and S velocities, $\alpha_{v},\alpha_{h},\beta_{v},\beta_{h}$.
Explicit relations can be found in appendix B. of \citet{SiLiTrTr07b}

### 2015-06-22 doc/USER_MANUAL/09_noise_simulations.tex

\subsection{Pre-simulation}
\begin{itemize}
\item As usual, we first configure the software package using:

\texttt{./configure FC=ifort MPIFC=mpif90}

Use the following if SCOTCH is needed:

\texttt{./configure FC=ifort MPIFC=mpif90 -{}-with-scotch-dir=/opt/scotch}\\

\item Next, we need to compile the source code using:

\texttt{make xgenerate\_databases}

\texttt{make xspecfem3D} \\

\item Before we can run noise simulations, we have to specify the noise
statistics, e.g., the ensemble-averaged noise spectrum. Matlab scripts
are provided to help you to generate the necessary file.

\texttt{EXAMPLES/noise\_tomography/NOISE\_TOMOGRAPHY.m} (main program)\\
 \texttt{EXAMPLES/noise\_tomography/PetersonNoiseModel.m}

In Matlab, simply run:

\texttt{NOISE\_TOMOGRAPHY(NSTEP, DT, Tmin, Tmax, NOISE\_MODEL)}\\

@@

After running the Matlab script, you will be given the following information
(plus a figure in Matlab):

\texttt{{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}{*}}\\
 \texttt{the source time function has been saved in:}\\
 \texttt{/data2/yangl/3D\_NOISE/S\_squared} (note this path must be
different)\\
 \texttt{S\_squared should be put into directory:}\\
 \texttt{OUTPUT\_FILES/NOISE\_TOMOGRAPHY/ in the SPECFEM3D Cartesian
package}\\

@@

As discussed in \citet{trompetal2010}, it takes three steps/simulations
to obtain one contribution of the ensemble sensitivity kernels:
\begin{itemize}
\item Step 1: simulation for generating wavefields\\
 \texttt{SIMULATION\_TYPE=1}\\
 \texttt{NOISE\_TOMOGRAPHY=1}\\
 \texttt{SAVE\_FORWARD} not used, can be either .true. or .false.
\item Step 2: simulation for ensemble forward wavefields\\
 \texttt{SIMULATION\_TYPE=1}\\
 \texttt{NOISE\_TOMOGRAPHY=2}\\
 \texttt{SAVE\_FORWARD=.true.}
\item Step 3: simulation for ensemble adjoint wavefields and sensitivity
kernels\\
 \texttt{SIMULATION\_TYPE=3}\\
 \texttt{NOISE\_TOMOGRAPHY=3}\\
 \texttt{SAVE\_FORWARD=.false.}

>>>>

\subsection{Pre-simulation}
\begin{itemize}
\item As usual, we first configure the software package using:
\begin{verbatim}
./configure FC=ifort MPIFC=mpif90
\end{verbatim}

Use the following if SCOTCH is needed:
\begin{verbatim}
./configure FC=ifort MPIFC=mpif90 --with-scotch-dir=/opt/scotch
\end{verbatim}

\item Next, we need to compile the source code using:
\begin{verbatim}
make xgenerate_databases
make xspecfem3D
\end{verbatim}

\item Before we can run noise simulations, we have to specify the noise
statistics, e.g., the ensemble-averaged noise spectrum. Matlab scripts
are provided to help you to generate the necessary file:
\texttt{EXAMPLES/noise\_tomography/NOISE\_TOMOGRAPHY.m} (main program)\\
\texttt{EXAMPLES/noise\_tomography/PetersonNoiseModel.m}

In Matlab, simply run:
\begin{verbatim}
NOISE_TOMOGRAPHY(NSTEP, DT, Tmin, Tmax, NOISE_MODEL)
\end{verbatim}

@@

After running the Matlab script, you will be given the following information
(plus a figure in Matlab):
\begin{verbatim}
*************************************************************
the source time function has been saved in:
..../S_squared (note this path must be different)
S_squared should be put into directory:
OUTPUT_FILES/NOISE_TOMOGRAPHY/ in the SPECFEM3D Cartesian package
\end{verbatim}

@@

As discussed in \citet{trompetal2010}, it takes three steps/simulations
to obtain one contribution of the ensemble sensitivity kernels:
\begin{itemize}
\item Step 1: simulation for generating wavefields
\begin{verbatim}
SIMULATION_TYPE = 1
NOISE_TOMOGRAPHY = 1
SAVE_FORWARD (not used, can be either .true. or .false.)
\end{verbatim}

\item Step 2: simulation for ensemble forward wavefields
\begin{verbatim}
SIMULATION_TYPE = 1
NOISE_TOMOGRAPHY = 2
SAVE_FORWARD = .true.
\end{verbatim}

\item Step 3: simulation for ensemble adjoint wavefields and sensitivity
kernels
\begin{verbatim}
SIMULATION_TYPE = 3
NOISE_TOMOGRAPHY = 3
SAVE_FORWARD = .false.
\end{verbatim}

### 2015-06-22 doc/USER_MANUAL/15_informations_for_developers.tex

You can get a very simple 1D version of a demo code (there is one in Fortran and one in Python):\\

git clone --recursive https://github.com/geodynamics/specfem1d.git\\

@@

\noindent For information on how to contribute to the code, i.e., for how to make your modifications, additions or improvements part of the
official package, see https://github.com/geodynamics/specfem3d/wiki .

>>>>

You can get a very simple 1D version of a demo code (there is one in Fortran and one in Python):
\begin{verbatim}
git clone --recursive https://github.com/geodynamics/specfem1d.git
\end{verbatim}

@@

\noindent For information on how to contribute to the code, i.e., for how to make your modifications, additions or improvements part of the
official package, see \url{https://github.com/geodynamics/specfem3d/wiki} .

### 2015-06-22 doc/USER_MANUAL/B_channel_codes.tex

Seismic networks, such as the Global Seismographic Network (GSN),
generally involve various types of instruments with different bandwidths,
sampling properties and component configurations. There are standards
to name channel codes depending on instrument properties. IRIS \texttt{(www.iris.edu)}
uses SEED format for channel codes, which are represented by three
letters, such as \texttt{LHN}, \texttt{BHZ}, etc. In older versions
of the SPECFEM3D Cartesian package, a common format was used for the
channel codes of all seismograms, which was \texttt{BHE/BHN/BHZ} for
three components. To avoid confusion when comparison are made to observed
data, we are now using the FDSN convention (\url{http://www.fdsn.org/})
for SEM seismograms. In the following, we give a brief explanation
of the FDSN convention used by IRIS, and how it is adopted in SEM
seismograms. Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm}
for further information.\\

@@

\caption{The band code convention is based on the sampling rate and the response
band of instruments. Please visit \texttt{www.iris.edu/manuals/SEED\_appA.htm}
for further information. Grey rows show the relative band-code range
in SPECFEM3D Cartesian, and the band codes used to name SEM seismograms
are denoted in red.}

>>>>

Seismic networks, such as the Global Seismographic Network (GSN),
generally involve various types of instruments with different bandwidths,
sampling properties and component configurations. There are standards
to name channel codes depending on instrument properties. IRIS (\url{http://www.iris.edu})
uses SEED format for channel codes, which are represented by three
letters, such as \texttt{LHN}, \texttt{BHZ}, etc. In older versions
of the SPECFEM3D Cartesian package, a common format was used for the
channel codes of all seismograms, which was \texttt{BHE/BHN/BHZ} for
three components. To avoid confusion when comparison are made to observed
data, we are now using the FDSN convention (\url{http://www.fdsn.org/})
for SEM seismograms. In the following, we give a brief explanation
of the FDSN convention used by IRIS, and how it is adopted in SEM
seismograms. Please visit \url{http://www.iris.edu/manuals/SEED_appA.htm}
for further information.\\

@@

\caption{The band code convention is based on the sampling rate and the response
band of instruments. Please visit \url{http://www.iris.edu/manuals/SEED_appA.htm}
for further information. Grey rows show the relative band-code range
in SPECFEM3D Cartesian, and the band codes used to name SEM seismograms
are denoted in red.}
